Fast and exact visibility on digitized shapes and application to saliency-aware normal estimation

### 2025-04-24 DGMM2025/paper.tex
Working on 4

    \newcommand{\Kernel}[1]{\ensuremath{w_{\sigma}(#1)}}
    \paragraph*{Saliency-aware normal estimator.}
    We propose a new normal estimator on digital surfaces, which uses
    the visibility of the point of interest while taking into account
    a user-given scale $\sigma$. If $\sigma$ is proportionnal to the
    average length between visible points (so some
    $\Theta(\sqrt{h})$), then this estimator is observed to be
    multigrid convergent along digitization of smooth shapes. However,
    since the computation window is limited by the visibility, it
    better approximates the geometry near sharp or salient features.

    More precisely, we choose a Gaussian function
    $\Kernel{x}\coloneqq e^{-\frac{x^2}{2\sigma^2}}$ as a weight kernel. Let
    $V_p$ be the set of visible points from the point $p$.  Let $c_p$
    be the weighted centroid of the visible points around $p$,
    i.e. $c_p \coloneqq \frac{\sum_{q \in V_p} \Kernel{\|q-p\|}q}{\sum_{q \in
        V_p} \Kernel{\|q-p\|}}$.  We form the weighted covariance matrix
    $\mathcal{V}_p$ of the points $V_p$ as:
    \begin{equation}
        \mathcal{V}_p = \sum_{q \in V_p} \Kernel{\|q-p\|}(q - c_p)(q - c_p)^T.
    \end{equation}
    The \emph{visibility normal} $\vec{n}_p$ of point $p$ is defined
    as the first eigenvector of the covariance matrix $\mathcal{V}_p$
    of the visible points, corresponding to its smallest
    eigenvalue. Its orientation is chosen so as to point in the same
    direction as the average of the trivial normals to the surfels
    touching the point $p$. Figure~\ref{fig:normals-estimation} shows an example
    of our normals compared to normals computed with the integral invariant (II) estimator~\cite{Lachaud:2017-lnm}.

    \begin{figure}
        \centering
        \begin{tabular}{|c||c|c|}
            \hline
            II normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-II} \\
            \hline
            Visibility normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-NV} \\
            \hline
        \end{tabular}
        \caption{Examples of normals computed on a chinese dragon as a color map. First set displays II normals, second set uses the visibility
        algorithm to compute them. The normals are colored according to their orientation as $0.5 \cdot (n_i + \mathbbm{1})$
            . The right images are the smooth version of the left images}
        \label{fig:normals-estimation}
    \end{figure}

    Then, to compute the curvatures, we use the Corrected Normal Current
    estimator~\cite{lachaud:2022-dcg} which are based on the normals computed
    above.

    \paragraph*{Application to curvature estimation.}
    In order to illustrate better that the visibility normals better
    take into account the sharp features of a digital surface while
    staying meaningfull in digitization of smooth parts, we compare
    the curvature estimates induced by different discrete normal
    estimators. We exploit the \emph{Corrected Normal Current (CNC)
    estimator}, whose theory is detailed in \cite{lachaud:2022-dcg}
    and which can estimate all kinds of curvatures given positions and
    normals. It produces state-of-the-art curvature estimates on
    polygonal surfaces \cite{lachaud:2020-cgf}, point clouds
    \cite{lachaud:2023-cgf}, and even outperforms Integral Invariant
    curvature estimates \cite{coeurjolly:2014-cviu} on digital
    surfaces \cite{lachaud:2022-dcg}. Its implementation is also
    available in    \href{https://dgtal-team.github.io/doc-nightly/moduleCurvatureMeasures.html}{\textsc{DGtal}}.

>>>>

    \newcommand{\Kernel}[1]{\ensuremath{w_{\sigma}(#1)}}
    \paragraph{Saliency-aware normal estimator.}
    We propose a new normal estimator on digital surfaces, which uses
    the visibility of the point of interest while taking into account
    a user-given scale $\sigma$. If $\sigma$ is proportionnal to the
    average length between visible points (so some
    $\Theta(\sqrt{h})$), then this estimator is observed to be
    multigrid convergent along digitization of smooth shapes. However,
    since the computation window is limited by the visibility, it
    better approximates the geometry near sharp or salient features.

    More precisely, we choose a Gaussian function
    $\Kernel{x}\coloneqq e^{-\frac{x^2}{2\sigma^2}}$ as a weight kernel. Let
    $V_p$ be the set of visible points from the point $p$.  Let $c_p$
    be the weighted centroid of the visible points around $p$,
    i.e. $c_p \coloneqq \frac{\sum_{q \in V_p} \Kernel{\|q-p\|}q}{\sum_{q \in
        V_p} \Kernel{\|q-p\|}}$.  We form the weighted covariance matrix
    $\mathcal{V}_p$ of the points $V_p$ as:
    \begin{equation}
        \mathcal{V}_p = \sum_{q \in V_p} \Kernel{\|q-p\|}(q - c_p)(q - c_p)^T.
    \end{equation}
    The \emph{visibility normal} $\vec{n}(p)$ of point $p$ at scale $\sigma$ is defined
    as the first eigenvector of the covariance matrix $\mathcal{V}_p$
    of the visible points, corresponding to its smallest
    eigenvalue. Its orientation is chosen so as to point in the same
    direction as the average of the trivial normals to the surfels
    touching the point $p$. Figure~\ref{fig:normals-estimation} shows an example
    of our normals compared to normals computed with the integral invariant (II) estimator~\cite{Lachaud:2017-lnm}.

    \begin{figure}
        \centering
        \begin{tabular}{|c||c|c|}
            \hline
            II normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-II} \\
            \hline
            Visibility normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-NV} \\
            \hline
        \end{tabular}
        \caption{Examples of normals computed on a chinese dragon as a
          color map. First set displays II normals, second set uses
          the visibility algorithm to compute them. The normals are
          colored according to their orientation as $0.5 \cdot (n_i +
          \mathbbm{1})$. The right images are the smooth version of
          the left images}
        \label{fig:normals-estimation}
    \end{figure}

    %% Then, to compute the curvatures, we use the Corrected Normal Current
    %% estimator~\cite{lachaud:2022-dcg} which are based on the normals computed
    %% above.

    \paragraph{Application to curvature estimation.}
    In order to illustrate better that the visibility normals better
    take into account the sharp features of a digital surface while
    staying meaningfull in digitization of smooth parts, we compare
    the curvature estimates induced by different discrete normal
    estimators. We exploit the \emph{Corrected Normal Current (CNC)
    estimator}, whose theory is detailed in \cite{lachaud:2022-dcg}
    and which can estimate all kinds of curvatures given positions and
    normals. It produces state-of-the-art curvature estimates on
    polygonal surfaces \cite{lachaud:2020-cgf}, point clouds
    \cite{lachaud:2023-cgf}, and even outperforms Integral Invariant
    curvature estimates \cite{coeurjolly:2014-cviu} on digital
    surfaces \cite{lachaud:2022-dcg}. Its implementation is also
    available in    \href{https://dgtal-team.github.io/doc-nightly/moduleCurvatureMeasures.html}{\textsc{DGtal}}.
